---
title: "spring_21"
author: "Selam Berekat"
date: "6/1/2021"
output: html_document
---

#Assignment:
#Correlation matrix for all key toxicity variables?  - Selam
#A second plot showing a relationship between the variable and another key variable of interest. - Selam

```{r, echo = FALSE}
# Load packages
library(tidyverse)
library(tidyr)
library(ggplot2)
library(corrplot)
```


```{r, echo = FALSE}
osha <- read.csv("dataset.csv", header = T)
y1 <- as.factor(osha$union_status)
x1 <- as.numeric(osha$gravity)
x2 <- as.factor(osha$insp_scope)
x3 <- as.factor(osha$insp_type)
x4 <- osha$noviol

x5 <- as.numeric(osha$current_penalty)
x6 <- as.numeric(osha$initial_penalty)
df <- data.frame(y1, x1, x2, x3, x4, x5, x6)
colnames(df) <- c("Union", "Gravity", "Ins-Scope", "Ins-Type", "Noviol", "Cur-Penalty", "Init-Penalty")

prop <- apply( apply(df, 2, is.na),  2, sum )/nrow(df) #proportion
prop

# (freq1 <- table(df$Union, df$Gravity, useNA="always"))
# (freq2 <- table(df$Union, df$`Ins-Scope`, useNA="always"))
# (freq3 <- table(df$Union, df$`Ins-Type`, useNA="always"))
# (freq4 <- table(df$Union, df$Noviol, useNA="always"))


# head(df, 10)
# df <- na.omit(df)
# summary(x4)
# sum(is.na(x4))
# which(is.na(df))
# fac <- factor(x4)
# levels(fac)
```
The variable "Gravity" has NA's that are worth 70% of our dataset and thus we decided to have it removed. Because it was affecting the variable "Noviol", for instance: for all the NA's on Gravity, they were a True for variable noviol, and removing the NA's were removing the entire True values in the variable Noviol.


# Correlation matrix
```{r}
df[] <- lapply(df, as.numeric)
corrplot(cor(df, use = "pairwise.complete.obs"))


library("PerformanceAnalytics")
library(EnvStats)
chart.Correlation(df, histogram = TRUE, pch = 19)
```

# Multiple Linear Regression Analysis Model
# Part (1)_using computational
# Coefficients
# Hypothesis Test
# Confidence Interval
# Choose best model meaning which model is best and what are the selected variables that can predict if there is relationship between "Penalty Fee and Type of Violation"


```{r}
df2 <- data.frame(y1, x2, x3, x4, x5, x6)
colnames(df2) <- c("Union", "InsScope", "InsType", "Noviol", "CurPenalty", "InitPenalty")
# df2 <- na.omit(df2)

##Y1: Union
df2 <- df2 %>% mutate(Union = ifelse((Union == "Y" | Union == "U" | Union == "A"), 1, 0))

##X4: Noviol
df2$Noviol = ifelse(df2$Noviol == "TRUE", 1, 0) #[ 1=True, No Viol and 0=False, Yes Viol ]

##X2: Inspection Scope
df2$InsScope <- factor(df2$InsScope, levels = c("A", "B", "D"),
                       labels = c("1", "2", "3"))

##X3: Inspection Type
df2$InsType <- factor(df2$InsType, levels = c("A", "B", "C", "F", "G", "H", "J", "M"),
                      labels = c("1", "2", "3", "4", "5", "6", "7", "8"))

# dim(df2$Union)
# dim(df2$InsType)
# table(df2$Union, df2$InspType)
table(df2$Noviol)
# head(df2)
# str(df2)
```

