---
title: "osha"
author: "Selam B"
date: "1/26/2021"
output: html_document
---

Research question: How do we examine toxicity? 
I will use 2 variables "number of employees exposed" and " facilities, sic_code". 
Since there are many facilities in CA, i decided to use the variable "sic_code" stands for standard industrial classification code. 

# Load packages

```{r, echo = FALSE}
library(lubridate)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)
library(questionr)
```

# Load dataset 

-removed NA
-there are 339 observation with 71 columns
-plot original data



```{r, echo = FALSE}
osha <- 
  read.csv("CA_OSHA_prison_insp_viol_2010_2019_cleaned_long.csv") %>% 
  filter(!is.na(nr_exposed) & !is.na(year) & !is.na(sic_code)) 

#Histogram
ggplot(osha, aes(x = nr_exposed, fill = as.factor(sic_code))) + geom_histogram(binwidth = 10, position="identity") + ggtitle("Distribution of Hazard Exposed Employees by", subtitle = "Type of Standard Industrial Classification") + theme_classic(base_size = 10) + scale_fill_brewer(palette = "Paired")

#Boxplot
ggplot(osha, aes(y = nr_exposed, x = as.factor(sic_code))) + geom_boxplot() + ylab("Miles per Gallon") + xlab("Standard Industrial Classification") + ggtitle("Distribution of Employee Hazard Exposure by", subtitle = "Facility Standard Industrial Classification Code")  
```


We observe a right skewed histogram


# Data exploration before data visualization

1. Organize the year in dataset according to open date of case for exposed employees.

```{r fig.height=6, fig.width=6, echo = FALSE}
# sort(unique(osha$nr_exposed))
# sort(unique(osha$sic_code))

# reformating the open dates
osha$open_date = ymd(osha$open_date)
osha$year = year(osha$open_date)

#test to check the new column have the correct year 
table(osha$year)
```

2. For the purpose to classify the number of employees by parts; 

I convert the numeric variable (nr_exposed) to categorical (expose) variables. 

```{r fig.height=10, fig.width=10, echo = FALSE}
#recode nr_exposed
osha$expose = cut(osha$nr_exposed, 
                        c(-1, 500, 1000, 2000, 3000),
                        dig.lab = 5) # stops scientific notation

osha$expose = factor(osha$expose,
                        labels = c("Less than 500", "500 - 1000", "1000 - 2000", "Above 2000"),
                        exclude = NA)

#test our new column is correct
two.way1 = table(osha$expose, osha$nr_exposed)

#Plot the data with the 2 new column
two.way2 = table(osha$sic_code, osha$expose)
barplot(two.way2, beside = TRUE, legend.text =  rownames(two.way2), main = "Distribution of Employee Hazard Exposure by Facility Standard Industrial Classification Code")
```


# Data Visualize 

# Facet by Year

```{r fig.height=10, fig.width=10, echo = FALSE}
ggplot(osha, aes(x = expose, fill = sic_code)) + 
  geom_bar(position = "dodge") +
  
  facet_grid(rows = vars(year)) +
  scale_y_continuous(labels = scales::comma) +
  coord_flip() + theme_minimal(base_size = 15) +
  labs(x = "Number of Exposed Employees", y = "Standard Industrial Classification Code", 
       title = "Distribution of Employee Hazard Exposure by", subtitle = "Facility Standard Industrial Classification Code by year 2000-2014")
```


# Classify the variable Region

```{r, echo = FALSE}
sort(unique(osha$site_city))
```

```{r, echo = FALSE}
regions <- read.csv("/Users/ksalamawit/Desktop/Github/Book.csv")
regions <- 
  regions %>%
  mutate(site_city = str_trim(site_city))

osha_joined <-
osha %>% 
  left_join(regions, by = "site_city")
```

# Data visualize 

# Facet by year and region "from 2010 to 2014"

```{r, echo = FALSE}
osha_joined %>% filter(year %in% 2010:2014) %>%
  ggplot(aes(x = expose, fill = region)) +
  geom_bar() + facet_grid(rows = vars(year)) +
  scale_y_continuous(labels = scales::comma) +
  coord_flip() + theme_minimal(base_size = 15) +
  labs(x = "Number of Exposed Employees", y = "Standard Industrial Classification Code", 
        title = "Distribution of Employee Hazard Exposure by Facility Standard Industrial Classification Code",
       subtitle = "Faceted by Year, 2010-2014") 
```


# Plot the data and facet by year and region "from 2000 to 2010"

```{r fig.height=10, fig.width=10, echo = FALSE}
osha_joined %>% 
  filter(year %in% 2000:2010) %>%
  ggplot(aes(x = nr_exposed, fill = region)) +
  geom_histogram(binwidth = 500) + 
  facet_grid(rows = vars(year)) + 
  scale_y_continuous(labels = scales::comma) +
  #coord_flip() + 
  theme_minimal(base_size = 15) +
  labs(x = "Number of Exposed Employees", y = "Facilities", 
       title = "Distribution of Employee Hazard Exposure by Facility Standard Industrial Classification Code",
       subtitle = "Faceted by Year, 2000-2010") 
```


# Plot using column graph (not correct)

```{r fig.height=6, fig.width=10, echo = FALSE}
osha_joined %>%
  group_by(region, year) %>%
  summarize(total_nr_exposed = sum(nr_exposed)) %>%
  ggplot(aes(x = year, y = total_nr_exposed, fill = region)) +
  geom_col()

#not correct - I don't know why
osha_joined %>%
  ggplot(aes(x = year, y = nr_exposed, fill = region)) +
  stat_summary(geom = "col", fun = "sum")
```